% Copyright (C) 2019 Esote.
% Permission is granted to copy, distribute and/or modify this document
% under the terms of the GNU Free Documentation License, Version 1.3
% or any later version published by the Free Software Foundation;
% with no Invariant Sections, no Front-Cover Texts, and no Back-Cover Texts.
% A copy of the license is included in the section entitled "GNU
% Free Documentation License".

\documentclass{beamer}
\mode<presentation> {
\usetheme{Madrid}
\setbeamertemplate{navigation symbols}{}
}

\title[BonesBSD\hspace{1em}\texttt{\href{https://github.com/esote/bonesbsd}{github.com/esote/bonesbsd}}]{Introducing BonesBSD}

\date{April 24, 2019}

\begin{document}

\begin{frame}
	\titlepage
	\centering
	\texttt{\href{https://github.com/esote/bonesbsd}{github.com/esote/bonesbsd}}
\end{frame}

\begin{frame}{Overview}
	\tableofcontents
\end{frame}

\section{About}

\setlength{\parskip}{3em}

\begin{frame}{About}
	BonesBSD is a fork of the OpenBSD kernel on \texttt{-current}.

	The goal is to maintain the most useful or critical functionality,
	removing unused parts of the kernel.
\end{frame}

\section{History}

\begin{frame}{History}
	\begin{table}
		\begin{tabular}{|c|c c|}
			\hline
			Change & Diff & \\
			\hline
			Initial commit & +3,174,432 & \\
			Remove macppc, socppc & +13 & -47,349 \\
			Remove powerpc & +15 & -12,039 \\
			Remove hppa & +13 & -53,093 \\
			Remove alpha & +26 & -63,034 \\
			Remove loongson, mips64, octeon, sgi & +15 & -129,373 \\
			Remove sparc64 and afb microcode & +27 & -125,641 \\
			\hline
		\end{tabular}
	\end{table}
\end{frame}

\begin{frame}{History}
	\begin{table}
		\begin{tabular}{|c|c c|}
			\hline
			Change & Diff & \\
			\hline
			Remove m88k, luna88k & +3 & -46,583 \\
			Remove landisk & +8 & -8,805 \\
			Remove arm & +14 & -90,010 \\
			Remove sh & & -20,274 \\
			Remove NTFS & +2 & -4,414 \\
			Remove NFS and \texttt{nfssvc(2)} & +22 & -18,064 \\
			Remove FUSE & +7 & -3,331 \\
			Remove drm (inteldrm, radeondrm) & +6 & -394,799 \\
			\hline
		\end{tabular}
	\end{table}
\end{frame}

\begin{frame}{History}
	\setlength{\parskip}{\baselineskip}

	Lines before: 3,174,432

	Lines after: 336,921 (89\% decrease)

	Compiled size before: 14.7M (amd64)

	Compiled size after: 10.9M (25\% decrease)

	Still functions the same on amd64 and i386.
\end{frame}

\section{Breaking Userspace}

\begin{frame}{Breaking Userspace}
	Breaking userspace is inevitable. If you want to use an architecture no
	longer supported: too bad.

	For cases of noticeable errors in userspace, if they're avoidable I
	generate a patch.

	Ex: \texttt{patches/0001-Remove-rc-NFS-reference.patch}
\end{frame}

\section{Installation}

\setlength{\parskip}{\baselineskip}

\begin{frame}{Installation}
	\begin{enumerate}
		\item Install OpenBSD
		\item Clone BonesBSD
		\item \texttt{cd bonesbsd/sys/arch/\$(machine)/compile/GENERIC.MP/}
		\item \texttt{make obj}
		\item \texttt{make config}
		\item \texttt{ln -sf /usr/bin/gcc /usr/bin/cc}
		\item \texttt{make} (until error occurs assembling \texttt{locore.S})
		\begin{itemize}
			\item Might care enough to actually fix this later
		\end{itemize}
		\item \texttt{ln -sf /usr/bin/clang /usr/bin/cc}
		\item Remove line 43 from \texttt{obj/Makefile} (\texttt{-msave-args})
		\item \texttt{make}
		\item \texttt{make install}
	\end{enumerate}
	Old kernel is moved to \texttt{/obsd}.
\end{frame}

\section{What's next}

\begin{frame}{What's next}
	Still plenty of things to remove:
	\begin{itemize}
		\item \textbf{ipv6} (ie \texttt{sys/netinet6/})
		\item IBM eFUSE
		\item isofs (ie \texttt{sys/isofs/}, both cd9660 and udf)
		\item Xenocara, aperture driver, etc. X.org remnants
		\item msdosfs (ie \texttt{sys/msdosfs/})
		\item Hyper-V virtualization (ie \texttt{sys/arch/\$(machine)/})
		\item VMware tools (ie \texttt{sys/arch/\$(machine)/})
	\end{itemize}
\end{frame}

\begin{frame}{What's next}
	Undecided:
	\begin{itemize}
		\item mouse? (pseudo-devices, wsmouse daemon, drivers, etc.)
		\item wireless? (IEEE 802.11, ie \texttt{sys/net80211/})
		\item audio? (and sndio) --- will likely require userspace patches
		\item fifofs? (ie \texttt{sys/miscfs/fifofs/})
		\item deadfs? (ie \texttt{sys/miscfs/deadfs/})
	\end{itemize}

	Complex:
	\begin{itemize}
		\item Merge upstream changes occasionally (maybe only every time OpenBSD has a release)
		\item Rewrite certain userspace things that break, so they don't depend on removed syscalls.
	\end{itemize}
\end{frame}

\end{document}
